---
title: "Deep Learning (Independent Study)"
author: "Saheed Adisa, Ganiyu"
date: "2023-10-20"
output:
  html_document:
    df_print: paged
---

# 7. Fit a neural network to the Default data. Use a single hidden layer with 10 units, and dropout regularization. Have a look at Labs 10.9.1–10.9.2 for guidance. Compare the classifcation performance of your model with that of linear logistic regression.


## Loading the Default dataset
```{r }
library(ISLR2)
default_data <- Default
head(default_data)
```



### Data summary and structure
```{r }
#summary(default_data)
str(default_data)
```


## Spliting the data
```{r }
test_indices <- sample(1:nrow(default_data), 0.25*nrow(default_data))

# Create the training and test datasets
train_data <- default_data[-test_indices, ]
test_data <- default_data[test_indices, ]
```


# =============== Logistic Regression =======================

## Fitting logistic regression
```{r }
logistic_model <- glm(default ~ ., data = train_data, family = binomial(link = "logit"))
logistic_training_probabilities <- predict(logistic_model, type = "response", newdata = test_data)
logistic_pred <- ifelse(logistic_training_probabilities > 0.5, "Yes", "No")
logistic_accuracy <- mean(logistic_pred == test_data$default)
cat("Logistic Regression Accuracy:", logistic_accuracy*100,"%")
```



# =================== Neural  network ========================

## Neural Network Visualization
```{r }
library(magrittr)  # for %<>% operator
library(dplyr)
library(neuralnet)
default_data2 <- Default

# converting factor variable into numeric
default_data2$default <- ifelse(default_data2$default == "Yes", 1, 0)
default_data2$student <- ifelse(default_data2$student == "Yes", 1, 0)

# creating neural model
neural_model <- neuralnet(default ~ .,
                          data = default_data2,
                          hidden = 10,
                          linear.output = F,
                          lifesign = "full",
                          rep = 1)
plot(neural_model,
     col.hidden = 'darkgreen',
     col.hidden.synapse = 'darkgreen',
     show.weights = F,
     information = F,
     fill = 'orange')

```



## Spliting the dataset and normalizing
```{r }
default_data2 <- as.matrix(default_data2)    # converting to matrix
train_data2 <- default_data2[-test_indices, 2:4]
test_data2 <- default_data2[test_indices, 2:4]
y_train2 <- default_data2[-test_indices, 1]
y_test2 <- default_data2[test_indices, 1]

# Normalizing
m <- colMeans((train_data2))
s<- apply(train_data2, 2, sd)
train_data2 <- scale(train_data2, center = m, scale = s)
test_data2 <- scale(test_data2, center = m, scale = s)
```



## Creating Neural model with keras package
```{r }
library(keras)
k_neural_model <- keras_model_sequential()
k_neural_model %>% 
  layer_dense(units = 10, activation = 'relu', input_shape = c(3)) %>%
  layer_dropout(rate = 0.4) %>%
  layer_dense(units = 1, activation = "sigmoid")

summary(k_neural_model)
```



## Compiling and fitting the model
```{r }
# Compile
k_neural_model%>% compile(loss = "binary_crossentropy",
optimizer = optimizer_rmsprop(),
metrics = c("accuracy"))

# Fitting
system.time(
history <- k_neural_model %>%
fit(train_data2, y_train2, epochs = 50, batch_size = 32,
validation_data = list(test_data2, y_test2)))

# Evaluate the model
eval <- k_neural_model %>% evaluate(test_data2, y_test2)
cat("Test accuracy:", eval[2], "\n")
plot(history, smooth = FALSE)
```


## Creating Neural model with keras package for tuning number of units and batch_sizes
```{r }
library(keras)

units_list <- c(5,10,15,20,25)
batch_size_list <- c(16,20,24,28,32,36,40)
accuracy_errors <- matrix(0, nrow = length(units_list), ncol = length(batch_size_list))

for (b in 1:length(batch_size_list)) {
  for (u in 1:length(units_list)) {
    library(keras)
    # Creating the model
    k_neural_model <- keras_model_sequential()
    k_neural_model %>% 
      layer_dense(units = u , activation = 'relu', input_shape = c(3)) %>%
      layer_dropout(rate = 0.3) %>%
      layer_dense(units = 1, activation = "sigmoid")

    # Compile
    k_neural_model%>% compile(loss = "binary_crossentropy",
    optimizer = optimizer_rmsprop(),
    metrics = c("accuracy"))

    # Fitting
    history <- k_neural_model %>%
    fit(train_data2, y_train2, epochs = 50, batch_size = 32,
    validation_data = list(test_data2, y_test2), verbose = 0)

    # Evaluate the model
    eval <- k_neural_model %>% evaluate(test_data2, y_test2)
    accuracy_errors[u, b] <- eval[2]
  }
}

columnname <- c("B_size=16", "B_size=20", "B_size=24", "B_size=28", "B_size=32"
                , "B_size=36", "B_size=40")
rowname <- c("units=5", "units=10" , "units=15", "units=20", "units=25")

colnames(accuracy_errors) <- columnname
rownames(accuracy_errors) <- rowname   
```

```{r }
accuracy_errors
```



## Creating Neural model with keras package with 2 layers
```{r }
 library(keras)
# Creating the model
k_neural_model <- keras_model_sequential()
k_neural_model %>% 
  layer_dense(units = 10, activation = 'relu', input_shape = c(3)) %>%
  layer_dense(units = 5, activation = 'relu') %>%
  layer_dropout(rate = 0.3) %>%
  layer_dense(units = 1, activation = "sigmoid")

summary(k_neural_model)

# Compile
k_neural_model%>% compile(loss = "binary_crossentropy",
optimizer = optimizer_rmsprop(),
metrics = c("accuracy"))

# Fitting
system.time(
history <- k_neural_model %>%
fit(train_data2, y_train2, epochs = 50, batch_size = 32,
validation_data = list(test_data2, y_test2), verbose = 0))

# Evaluate the model
eval <- k_neural_model %>% evaluate(test_data2, y_test2)
cat("Test accuracy:", eval[2], "\n")
plot(history, smooth = FALSE)

```

**Comment:** We observed that Logistic regression with $97.36 \%$ accuracy slightly perform better of Neural network of $97.68\%$ accuracy, but when added another layer with 5 neurons, we achieved the same accuracy with Logistic Regression. The 3 layer we later created performed a bit lower to others with $96.64\%$ accuracy.

**Observation:** softmax activation function perform poorly in output layer with $32\%$ accuracy.





## 5. Download the zip code data set from the book homepage. For both training and test sets, you need to unzip the files first and then load them into $\mathrm{R}$. Consider the 10-class classification problems with $Y \in\{0,1, \cdots, 9\}$. Apply LDA, the knn with $k=1,3,5,7,15$, respectively, and CNN. Report the training and test errors for LDA and $\mathrm{knn}$ with each choice of $k$. Summarize your findings.

# fitting LDA model
```{r }
library(MASS)  # for LDA
# Loading the dataset
train_data <- read.table("zip.train", sep = "")
test_data <- read.table("zip.test", sep = "")

# Function to calculate the error rate
calculate_error <- function(true_labels, predicted_labels) {
  return (mean(true_labels != predicted_labels))
}

lda_model <- lda(V1 ~ ., data = train_data)
lda_train_pred <- predict(lda_model, train_data)$class
lda_test_pred <- predict(lda_model, test_data)$class
lda_train_error <- calculate_error(train_data$V1, lda_train_pred)
lda_test_error <- calculate_error(test_data$V1, lda_test_pred)
cat("================== Training and testing errors ==================\n")
cat(paste("LDA Train Error:", lda_train_error, "\t LDA Test Error:", lda_test_error, "\n"))

```




# fitting k-NN classification for part
```{r }
library(class)  # for k-NN
k_values <- c(1, 3, 5, 7, 15)
cat("================== Training and testing errors ==================\n")
for (k in k_values) {
  knn_train_pred <- knn(train_data[, -1], train_data[, -1], train_data$V1, k)
  knn_train_error <- calculate_error(train_data$V1, knn_train_pred)
  knn_test_pred <- knn(train_data[, -1], test_data[, -1], train_data$V1, k)
  knn_test_error <- calculate_error(test_data$V1, knn_test_pred)
  cat(paste("\n k =", k, "\t k-NN Train Error:", knn_train_error, "\t k-NN Test Error:", knn_test_error, "\n"))
}

```



# =============== Fitting CNN ============================
```{r }
library(keras)
x_train_data <-as.matrix( train_data[,-1])
y_train_data <-train_data[,1]
x_test_data <-as.matrix(test_data[,-1])
y_test_data <-test_data[,1]
dimnames(x_train_data) <- list(NULL, NULL)
dimnames(x_test_data) <- list(NULL, NULL)
x_train_data <- array_reshape(x_train_data, dim = c(nrow(x_train_data), 16, 16, 1))
x_test_data <- array_reshape(x_test_data, dim = c(nrow(x_test_data), 16, 16, 1))


# one-hot encoding
y_train_data2 <- to_categorical(y_train_data, 10)
y_test_data2 <- to_categorical(y_test_data, 10)

# creating the CNN network
model <- keras_model_sequential() %>%
  layer_conv_2d(filters = 32, kernel_size = c(3, 3),
                padding = "same", activation = "relu",
                input_shape = c(16, 16, 1)) %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_conv_2d(filters = 64, kernel_size = c(3, 3),
                padding = "same", activation = "relu") %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_conv_2d(filters = 128, kernel_size = c(3, 3),
                padding = "same", activation = "relu") %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_conv_2d(filters = 256, kernel_size = c(3, 3),
                padding = "same", activation = "relu") %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_flatten() %>%
  layer_dropout(rate = 0.5) %>%
  layer_dense(units = 64, activation = "relu") %>%
  layer_dense(units = 10, activation = "softmax")
summary(model)
```




# Train the model
```{r}
model %>% compile(loss = "categorical_crossentropy",
optimizer = optimizer_rmsprop(), metrics = c("accuracy"))
history <- model %>% fit(x_train_data, y_train_data2, epochs = 30,
                         batch_size = 64, validation_split = 0.2)
test_pred<- k_argmax(model %>% predict(x_test_data)) 
test_Error_CNN <- mean(drop(y_test_data)!=as.numeric(test_pred)) 
cat("\n Test Error=", test_Error_CNN)
```

**Comment:** We observed that the CNN out perform bot LDA and KNN on the given zip code dataset.








# 8. From your collection of personal photographs, pick 10 images of animals (such as dogs, cats, birds, farm animals, etc.). If the subject does not occupy a reasonable part of the image, then crop the image. Now use a pretrained image classifcation CNN as in Lab 10.9.4 to predict the class of each of your images, and report the probabilities for the top fve predicted classes for each image.

## Loading the images and prepocessing
```{r }
library(jpeg)  # OR library(png)
library(keras)
img_dir <- "book_image"
image_names <- list.files(img_dir)
num_images <- length(image_names)
x <- array(dim = c(num_images, 224, 224, 3))
for (i in 1:num_images) {img_path <- paste(img_dir,
                                           image_names[i], sep = "/")
  img <- image_load(img_path, target_size = c(224, 224))
  x[i,,, ] <- image_to_array(img)
} 
x <- imagenet_preprocess_input(x)
```





# Viewing the images
```{r}
library(jpeg)  # OR library(png)

# List the image files in the directory
image_files <- list.files(img_dir, pattern = "\\.(jpg|png|jpeg|gif|bmp)$", full.names = TRUE)

# Take the first 10 image files 
image_files <- head(image_files, 8)

# Define the dimensions for each image
image_width <- 6  # Adjust as needed
image_height <- 6  # Adjust as needed

# Create a blank canvas for the grid
par(mfrow=c(2, 4), mar = c(0, 0, 0, 0))

# Loop through the images and display them
for (i in 1:8) {
  img <- readJPEG(image_files[i])  # Use readJPEG for JPEG images or readPNG for PNG images
  plot(0, 0, type = "n", axes = FALSE, xlab = "", ylab = "", xlim = c(0, image_width), ylim = c(0, image_height))
  rasterImage(img, 0, 0, image_width, image_height)
}

# Reset the plot layout
par(mfrow=c(1, 1))

```



## Fitting the model using pretrained model from resnet50
```{r }
CNN_model_pre <- application_resnet50(weights = "imagenet")
#summary(CNN_model_pre)
```



## Predicting the 10 images
```{r }
pred6 <- CNN_model_pre %>% predict(x) %>%
imagenet_decode_predictions(top = 3)
names(pred6) <- image_names
print(pred6)
```




```{r }

```



#==== Image classification using CNN model ======


```{r }
# Load Packages
library(EBImage)
library(keras)

# Read images
setwd('C:/Users/sahee/OneDrive/Desktop/Fall_2023/Ind_Study_STAT599/Helen_Rec/image15')
pics <- c('p1.jpg', 'p2.jpg', 'p3.jpg', 'p4.jpg', 'p5.jpg', 'p6.jpg',
          'c1.jpg', 'c2.jpg', 'c3.jpg', 'c4.jpg', 'c5.jpg', 'c6.jpg')
mypic <- list()
for (i in 1:12) {mypic[[i]] <- readImage(pics[i])}
```


### Vissualization
```{r }
display(mypic[[8]])
summary(mypic[[1]])
hist(mypic[[2]])
str(mypic)
```



### Preprocessing
```{r }
# Resize
for (i in 1:12) {mypic[[i]] <- resize(mypic[[i]],28,28)}

# Reshape
for (i in 1:12) {mypic[[i]] <- array_reshape(mypic[[i]], c(28, 28,3))}

# Row Bind
trainx <- NULL
for (i in 1:5) {trainx <- rbind(trainx, mypic[[i]])}
for (i in 7:11) {trainx <- rbind(trainx, mypic[[i]])}
str(trainx)
testx <- rbind(mypic[[6]], mypic[[12]])
trainy <- c(0,0,0,0,0,1,1,1,1,1 )
testx <- rbind(mypic[[6]], mypic[[12]])
testy <- c(0,1)

# One Hot Encoding
trainLabels <- to_categorical(trainy)
testLabels <- to_categorical(testy)
```




### Creating and compiling the model
```{r }
# Model
model <- keras_model_sequential()
model %>%
         layer_dense(units = 256, activation = 'relu', input_shape = c(2352)) %>%
         layer_dense(units = 128, activation = 'relu') %>%
         layer_dense(units = 2, activation = 'softmax')
summary(model)

# Compile
model %>%
         compile(loss = 'binary_crossentropy',
                 optimizer = optimizer_rmsprop(),
                 metrics = c('accuracy'))
```




### Fitting the model
```{r }
# Fit Model
history <- model %>%
         fit(trainx,
             trainLabels,
             epochs = 30,
             batch_size = 32,
             validation_split = 0.2)
```



### Evaluation
```{r }
# Evaluation & Prediction - train data
model %>% evaluate(trainx, trainLabels)
pred <- as.numeric(k_argmax(model %>% predict(trainx)))
table(Predicted = pred, Actual = trainy)
prob <- model %>% predict(trainx)
cbind(prob, Prected = pred, Actual= trainy)
```



```{r }

```














